---
title: "Supplementary Figure 12"
author: "Damian Wollny"
output:
  html_document:
    theme: paper
    df_print: kable
date: '`r format(Sys.Date(), "%B %d, %Y")`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = FALSE, message = FALSE)
```

```{r}
# Setup
library(tidyverse)
library(ggseqlogo)
library(cowplot)

theme_set(theme_classic(base_size=24))
```

<style type="text/css">

body, td {
   font-size: 18px;
}
</style>

## Figure S12a

Data was downloaded from <https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE99304> for <u>stress granule data</u> and from the Suppl. Data file `1-s2.0-S1097276517306512-mmc3.xlsx` for <u>p-body data</u>.
MEME was run on stress granule/p-body enriched transcripts as done for (e.g.) Fig.3b.
Prior to this script: run 'meme_parser_prob_mat.py' on MEME output to get input files in the right format for the ggseqlogo package

```{r, fig.height=12}
# great vignette for ggseqlogo found on: https://omarwagih.github.io/ggseqlogo/
# function to format parsed meme output
get_in_shape <- function(file){
  the_data <- read.table(file = file, sep = " ", skip = 1)
  mat <- rbind(the_data$V2,the_data$V4, the_data$V6, the_data$V8)
  rownames(mat) <- c("A", "C", "G", "T")
  return(mat)
}

# load all motif_*.txt files and apply the get_in_shape function
mat_list <- lapply(Sys.glob("./_khong/motif_*.txt"), get_in_shape)

# correct mat_list order and rename
mat_list <- mat_list[c(1,3,4,5,6,7,8,9,10,2)]
names(mat_list) <- paste("Motif:", c(1:10))

# Generate sequence logo
stress_gran_motifs <- ggseqlogo(mat_list, method = "bits", font = "helvetica_regular", ncol = 1) + theme(strip.text.x = element_text(size=22)) 
stress_gran_motifs
```

### Compare PDDA motif #1 to stress granule motif #10

```{r}

pdda_m1 <- get_in_shape("./_pdda/motif_1.txt")

sg_m10 <- get_in_shape("./_khong/motif_10.txt")

pdda_m1_plot <- ggseqlogo(pdda_m1, method = "bits", font = "helvetica_regular", ncol = 1) +
  annotate('rect', xmin = 10.5, xmax = 48.5, ymin = -0.05, ymax = 2.2, alpha = .1, col='black', fill='yellow') +
  theme(title = element_text(size=16)) +
  labs(title = "PDDA condensates: Motif #1")

sg_m10_plot <- ggseqlogo(sg_m10, method = "bits", font = "helvetica_regular", ncol = 1) +
  annotate('rect', xmin = 0.5, xmax = 38.5, ymin = -0.05, ymax = 2.2, alpha = .1, col='black', fill='yellow') +
  theme(title = element_text(size=16)) +
  labs(title = "Stress granules: Motif #10")

comp_sg <- plot_grid(pdda_m1_plot,sg_m10_plot, ncol = 1)
plot(comp_sg)
```

### Motifs enriched in P-body granule transcripts

```{r, fig.height=12}
# load all motif_*.txt files and apply the get_in_shape function
mat_list <- lapply(Sys.glob("./_hubstenberger/motif_*.txt"), get_in_shape)

# correct mat_list order and rename
mat_list <- mat_list[c(1,3,4,5,6,7,8,9,10,2)]
names(mat_list) <- paste("Motif:", c(1:10))

# Generate sequence logo
p_body_motifs <- ggseqlogo(mat_list, method = "bits", font = "helvetica_regular", ncol = 1) + theme(strip.text.x = element_text(size=22))
p_body_motifs
```

### Compare PDDA motif #1 to P-body motif #10

```{r}
pdda_m2 <- get_in_shape("./_pdda/motif_2.txt")

pb_m2 <- get_in_shape("./_hubstenberger/motif_2.txt")

pdda_m2_plot <- ggseqlogo(pdda_m2, method = "bits", font = "helvetica_regular", ncol = 1) +
  annotate('rect', xmin = 0.5, xmax = 40.5, ymin = -0.05, ymax = 2.2, alpha = .1, col='black', fill='yellow') +
  theme(title = element_text(size=16)) +
  labs(title = "PDDA condensates: Motif #2")

pb_m2_plot <- ggseqlogo(pb_m2, method = "bits", font = "helvetica_regular", ncol = 1) +
  annotate('rect', xmin = 10.5, xmax = 50.5, ymin = -0.05, ymax = 2.2, alpha = .1, col='black', fill='yellow') +
  theme(title = element_text(size=16)) +
  labs(title = "P-bodys: Motif #2")

comp_pb <- plot_grid(pdda_m2_plot, pb_m2_plot, ncol = 1)
plot(comp_pb)
```

