---
title: "Suppl. Figure 3 - Draft"
author: "Damian Wollny"
output:
  html_document:
    theme: paper
    df_print: kable
date: '`r format(Sys.Date(), "%B %d, %Y")`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = FALSE, message = FALSE)
options(width = 12)
```

```{r}
# Setup
library(tidyverse)
library(Rmpfr)
library(foreach)
library(data.table)
library(flowCore)
library(cowplot)

theme_set(theme_classic(base_size=24))
setwd(dir = "~/PD/Data/OoL/43_Draft_5/_s_fig9/")
```

<style type="text/css">

body, td {
   font-size: 18px;
}
</style>

### Motif Distances: Motif 1 & Motif 2

If they indeed form hairpins we would expect that the distances between motif 1 and motif 2 should be short. Note that each motif can be present several times on the same transcript. Hence, for each motif 1, I looked for <u>the closest</u> motif 2 since this would be the most probable candidate for a loop.

```{r}
# load and format input
input <- read.csv("/Users/damian_wollny/PD/Data/OoL/19.1_Draft_2/_Fig_3/_motif_distance_analysis/parsed_meme_output.csv", colClasses = c("motif_enrichment" = "character"))
input$motif_enrichment <- as.numeric(log(mpfr(input$motif_enrichment)))
input$motif_enrichment <- input$motif_enrichment*-1
input$motif_idx <- as.integer(input$motif)

# function from Ben Vernot to get combination of clostest motifs and their distances 
# needs data.table as input
get_combs <- function(dt) {
  res2 <- foreach (idx = 1:nrow(dt), .combine = rbind) %do% {
    t1 <- dt[idx, transcript]
    p1 <- dt[idx, start]
    m1 <- dt[idx, motif]
    
    #cat(idx, t1, m1, '\n')
    
    dt.tmp <- dt[transcript == t1 & motif != m1]
    if (nrow(dt.tmp) == 0) return(data.table())
    
    dt.tmp[, my.dist := abs(start - p1)]
    dt.tmp <- dt.tmp[my.dist == min(my.dist)]
    data.table(idx, t1, m1, m2 = dt.tmp[, motif], my.dist = dt.tmp[, my.dist])
  }
  res2
}

# dplyr::filter out only motif1 and motif2
input$motif <- as.character(input$motif)
dt <- input %>% dplyr::filter(motif == "ATYCTCCYGCCTCAGCCTCCCRARKWGCTGGGATTACAGGCGTGMGCCAC" |
                  motif == "GTGGCTCAYGCCTGTAATCCCAGCWMYTTGGGAGGCYGAGGCRGGAGRAT")
# sanity check
levels(as.factor(dt$motif))

# get motif distances
dist_dt <- get_combs(as.data.table(dt))

# plot distances
fig_s3a <- dist_dt %>% 
  ggplot(aes(my.dist)) +
  geom_histogram(color = "white") +
  labs(x = "Shortest distance between Motif1 & Motif2 [bp]")

plot(fig_s3a)

ggsave(fig_s3a, file="./fig3a.pdf", width = 8, height = 6)
```

```{r, fig.width=8}
dir <- "/Users/damian_wollny/PD/Data/OoL/18.2_integrative_analysis/_6_structure_analysis"

# import parsed RNAfold output for each transcript pool
enriched <- read.csv(file = paste(dir, "/output/enriched_clean.txt", sep = ""))
random <- read.csv(file = paste(dir, "/output/random_clean.txt", sep = ""))

rna_fold_results <- bind_rows(enriched, random) %>% 
  dplyr::rename(target_id = "transcript_id")

master <- readRDS(file = "/Users/damian_wollny/PD/Data/OoL/19.1_Draft_2/Generate_input/draft2_input.RDS")

rna_fold_results2 <- master %>% 
  inner_join(.,rna_fold_results, by = "target_id") %>% 
  dplyr::filter(experiment == "ool12") %>% 
  dplyr::select(c(target_id, log2_tpm,free_energy,length,transcript_pool)) %>% 
  mutate(norm_free_energy = free_energy/length)

fig_s3b <- ggplot(data = rna_fold_results2, aes(norm_free_energy))+
  geom_density(aes(colour = transcript_pool)) +
  scale_color_manual(values = c("#9E0142", "darkgrey"), 
                     labels = c("Enriched transcripts", "Random transcripts")) +
  xlab("free energy [kcal/mol] / length [bp]") +
  theme(legend.title = element_blank(), legend.position = "top")

plot(fig_s3b)

ggsave(fig_s3b, file="./fig3b.pdf", width = 7, height = 6)
```

```{r, fig.height=8, fig.width=10}
### Motif ###
#############
motif_30 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_motif_30min_015.fcs")
motif_60 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_motif_60min_016.fcs")

rc_motif_30 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_RC_motif_30min_017.fcs")
rc_motif_60 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_RC_motif_60min_018.fcs")

ds_motif_30 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_Motif^2_30min_023.fcs")
ds_motif_60 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_Motif^2_60min_024.fcs")

motif_60 <- as.data.frame(motif_60@exprs) %>% 
  mutate(group = "motif") %>% 
  mutate(oligo = "Motif1") %>% 
  mutate(ds = "no") %>% 
  mutate(time = "60")

rc_motif_60 <- as.data.frame(rc_motif_60@exprs) %>% 
  mutate(group = "motif") %>% 
  mutate(oligo = "RC of Motif1\n(Motif2)") %>% 
  mutate(ds = "no") %>% 
  mutate(time = "60")

ds_motif_60 <- as.data.frame(ds_motif_60@exprs) %>% 
  mutate(group = "motif") %>% 
  mutate(oligo = "Motif1 + RC") %>% 
  mutate(ds = "yes") %>% 
  mutate(time = "60")

# -------------------------------------------------------------------------------------------------------------------

### Scrambled motif ###
#######################
scrambled_30 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_SCR_30min_011.fcs")
scrambled_60 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_SCR_60min_012.fcs")

rc_scrambled_30 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_RC_SCR_30min_013.fcs")
rc_scrambled_60 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_RC_SCR_60min_014.fcs")

ds_scrambled_30 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_SCR^2_30min_021.fcs")
ds_scrambled_60 <- read.FCS(filename = "/Users/damian_wollny/PD/Data/OoL/21_Synthetic_oligo_analysis/20190731_FACS/20190731 Origin of life - Analyses_SCR^2_60min_022.fcs")


scrambled_60 <- as.data.frame(scrambled_60@exprs) %>% 
  mutate(group = "scrambled") %>% 
  mutate(oligo = "Scr Motif1") %>% 
  mutate(ds = "no") %>% 
  mutate(time = "60")


rc_scrambled_60 <- as.data.frame(rc_scrambled_60@exprs) %>%
  mutate(group = "scrambled") %>% 
  mutate(oligo = "RC of Scr Motif1") %>% 
  mutate(ds = "no") %>% 
  mutate(time = "60")


ds_scrambled_60 <- as.data.frame(ds_scrambled_60@exprs) %>% 
  mutate(group = "scrambled") %>% 
  mutate(oligo = "Scr Motif1 + RC") %>% 
  mutate(ds = "yes") %>% 
  mutate(time = "60")

### Generate master df ###
##########################
ool21_master <- bind_rows(motif_60, rc_motif_60, ds_motif_60, scrambled_60, rc_scrambled_60, ds_scrambled_60)

# convert 'group' col to factor and change level for facet labels
ool21_master$group <- as.factor(ool21_master$group)
levels(ool21_master$group) <- c("Motifs", "Scrambled")

fig_s3c <- ool21_master %>% 
  mutate(norm_fitc = `FITC-A`/`FSC-A`) %>% 
  dplyr::filter(time == 60) %>% 
  ggplot(aes(x =reorder(oligo, -norm_fitc, FUN = median), y = norm_fitc, colour = ds)) +
  geom_boxplot() +
  facet_grid(. ~ group,scales = "free_x") +
  scale_color_manual(values = c("#9E0142", "darkgrey"), 
                     breaks = c("yes", "no"),
                     labels = c("Double stranded", "Single stranded")) +
  geom_hline(aes(yintercept= median(ool21_master %>% 
                                      dplyr::filter(group == "Motifs", ds == "yes") %>% 
                                      mutate(norm_fitc = `FITC-A`/`FSC-A`) %>% 
                                      pull(norm_fitc))), colour="black", linetype="dashed") +
  labs(y = "normalized FAM signal [AU]") +
  theme(legend.title = element_blank(), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "top")

plot(fig_s3c)

ggsave(fig_s3c, file="./fig3c.pdf", width = 9, height = 7)
```

### Are the results significant?

```{r}
# format table
stat_test <- ool21_master %>% 
  mutate(norm_fitc = `FITC-A`/`FSC-A`) %>% 
  dplyr::filter(time == 60) %>% 
  select(oligo, norm_fitc) 

# pairwise test - bonferoni corrected
pairwise.t.test(stat_test$norm_fitc, stat_test$oligo, p.adjust.method = "bonf")
```

